---
title: "Reproducible"
author: "Zephyr Goriely"
date: "2023-12-02"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r loading in the data and packages, echo = FALSE, message = FALSE}

options(repos = c(CRAN = "https://cran.r-project.org/"))

install.packages("ggplot2")
install.packages("dplyr")
install.packages("palmerpenguins")
install.packages("janitor")

library(ggplot2)
library(dplyr)
library(palmerpenguins)
library(janitor)

source("functions/cleaning.r")

```

```{r bad figure code, echo=FALSE, message = FALSE}

penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows()

ggplot(data = penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, colour = species)) +
  scale_colour_manual(values = c("#21AAF3","blue", "#1C92D0")) +
   geom_smooth(method = "lm", se = FALSE, color = "cyan", linewidth = 2) +
  geom_point(size = 5) +
  scale_x_continuous(limits = c(2000, 7000)) +
  scale_y_continuous(limits = c(180, 220)) +
  labs(title = "Graph", x = "Flipper Length (m)", y = "Body Mass (kg)") +
  theme(legend.position = "none")

```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words).

## Here I have plotted a graph from the palmerpenguins dataset, comparing flipper length against body mass to investigate their linear relationship. However, the design choices I have made to plot the graph have resulted in a figure that is both misleading and confusing. Firstly, the graph is not appropriately labelled. There is no title and the axes labels are the wrong way around with the incorrect units. The axes should state: x = Body Mass (g), and y = Flipper lengh (mm), because of this, the displayed graph is a complete misrepresentation of the true relationship. The points on the scatter plot are enlarged such that the points overlap, reducing the clarity of the data spread and misrepresenting the number of points it also prevents the reader from identifying the exact position of each point. Although the points are colour coordinated to species, not only are the colours too similar to distinguish, but there is also no legend so the reader could not make use of this information. The scaling may be the biggest error in design choice. The axes are scaled very differently with the x-axis spanning 2000-7000 and y-axis between 180-220. The axes breaks are mis-representative of this difference in scale which makes the relationship appear much stronger than it may actually be. The y-axis is bounded such that data is cropped from either extreme and forced onto the limits. By forcing the data into these bounds, a relationship is manually forced and misleads the reader. The regression line plotted on the graph is embedded behind the data points and is essentially not visible, seeing as this graph is set to investigate the linear relationship, with the regression not visible and no provided information on the correlation coefficients, no satisfactory conclusions can be drawn.

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps, the figures visible, as well as clear code.*

*Your code should include the steps practiced in the lab session:*

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

-   ***New**: Run a statistical test*

-   ***New**: Create a Results Figure*

-   *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

### Introduction

```{r Data Exploration}

# The first step in data exploration is to install and load all the relevant packages. These packages include ggplot2 (to produce graphs), janitor (to help clean the dataset), dplyr (which allows us to use the pipe function), and palmerpenguins (which provides the dataset). This is done below:

install.packages("ggplot2")
install.packages("dplyr")
install.packages("palmerpenguins")
install.packages("janitor")

library(ggplot2)
library(dplyr)
library(palmerpenguins)
library(janitor)

# The next step is to investigate the raw data set that is already termed penguins_raw from the palmerpenguins package. By using the head() function, the dataset can be observed such that the aspects that require cleaning can be identified. 

head(penguins_raw)

# From the output, there are multiple noticable aspects of this dataset that need to be cleaned before further analysis. These are:
# The column names need to be changed so they are machine readable (lower and snake case) - this is done through the function 'clean_column_names'. 
# Any empty column or row should be removed - this is done through the function remove_empty_column_names. 
# The final three columns are not relevant and should be removed from the dataset - this is done through the functions select()
# The species names should be shortened - this is done through the function shorten_species()

# All of these cleaning functions are done below through a pipeline, defining a new object 'penguins_clean' which will be the newly cleaned dataset:

penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows() %>%
    select(-starts_with("Delta")) %>%
    select(-comments)
 
head(penguins_clean)

# Now that the dataset has been cleaned, a graph can be plotted to observe the data. Using ggplot, a scatterplot can be made to show the relationship between flipper length and body mass. The code for this is below:

ggplot(data = penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, colour = species)) +
  geom_point() +
  labs(title = "Body Mass plotted against Flipper length in all penguins", x = "Body Mass (g)", y = "Flipper Length (mm)") +
  geom_smooth(method = "lm", se = TRUE, aes(group = 1), color = "black") +
  theme_bw()
```

### Hypothesis

Based on the appearance of the exploratory graph and the plotted line of best fit, I formulate the hypothesis that flipper length increases with body mass, in other words, body mass explains some of the variation in flipper. This hypothesis supported by the apparent directly proportional relationship. This can be tested by investigating if there is a positive linear relationship between these variables. As such the hypothesis follows: Null hypothesis (H0): R\^2 = 0 - A coefficient of determination = 0 would suggest that body mass explains none of the variation in flipper length. Alternative hypothesis (H1): R\^2 \> 0 - A coefficient of determination \> 0 would suggest that body mass explains a non-zero amont of the variation in flipper length.

### Statistical Methods

```{r Statistics}

# The appropriate statistical test to investigate this relationship would be finding the coefficient of determination (R^2) in a linear model between these two variables. An F-test of overall significance can then be used to determine whether the found relationship is truly significant. All of these tests are conducted through the linear model function (lm()), 

linear_model <- lm(data = penguins_clean, flipper_length_mm ~ body_mass_g)
summary(linear_model)

x <- penguins_clean$body_mass_g
y <- penguins_clean$flipper_length_mm
correlation <- cor.test(x, y)
print(correlation)
```

### Results & Discussion

```{r Plotting Results}

# The summary table from that stats test provides a lot of information on the plotted relationship. However, for the sake of this hypothesis, the most important value is the adjusted R-squared. The adjusted R-sqaured value accounts for any overfitting from the regular R-squared. R-squared is a measure of how well the independent variable in the linear regression explains the variability of the dependent variable. In this context, 'Adjusted R-squared' is used to measure how much of the variation in flipper length is explained by body mass. R-squared values are bound between 0 and 1, higher values indicate stronger relationships. 

# The results of the stat test produce an Adjusted R-squared value of 0.7583. The high value demonstrates a very strong relationship, suggesting that body mass explains a lot of the variation in flipper length. Additionally, the significance of this value is supported by the provided F-test which produces a p-value < 2.2e-16 which is much smaller than 0.05. This allows us to reject the null hypothesis and claim that the relationship is not likely due to chance.

# For full visibility, the initial graph is plotted again, this time only focussing on the tested relationship. The results of the stats test are also projected. The code for the graph is shown below:

ggplot(data = penguins_clean, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point() +
  labs(title = "Body Mass plotted against Flipper length in all penguins", x = "Body Mass (g)", y = "Flipper Length (mm)") +
  geom_smooth(method = "lm", se = TRUE, aes(group = 1), color = "black") +
  theme_bw() +
  geom_label(aes(x = 5000, y = 190), hjust = 0, label = paste("Adj R2 = ",signif(summary(linear_model)$adj.r.squared, 5),
                                               "\nIntercept =",signif(linear_model$coef[[1]],5 ),
                                               " \nSlope =",signif(linear_model$coef[[2]], 5),
                                               " \nP =",signif(summary(linear_model)$coef[2,4], 5)))


```

### Conclusion

## This leads to the conclusion that body mass explains variation in flipper length and those penguins with larger body mass are more likely to have longer flippers. This conclusion would be expected as greater body masses usually indicate larger birds and given that both of these measures are of morphological features, it makes sense that flipper length will increase as body mass increases.

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
